Feasibility of a Cluster Randomised Trial on the Effect of Trauma Life Support Training: A Pilot Study in India

### 2025-09-01 manuscript.Rmd
Finalise revising the manuscript

Trauma, defined as a clinical entity composed of physical injury and the associated response of the body, causes 4.3 million deaths every year [@gbd_2021_collaborators_injuries_2024]. Several trauma life support training programmes have been developed to improve the early management of patients in hospitals by providing a structured framework for assessment and treatment [@Mohammad2013; @Jayaraman2014; @Kadhum2020].

@@

There are no controlled trials of the effect of ATLS or PTC on patient outcomes, and no trials evaluating any trauma life support training programme in a general trauma population [@Mohammad2013; @Jayaraman2014; @Kadhum2020; @Jin2021; @putra_impact_2023; @mciver_effect_2024], although a recent cluster randomised trial found that the rural trauma team development course reduces mortality in patients with motorcycle injuries [@lule_impact_2025]. Systematic reviews call for controlled trials of ATLS and PTC in settings where these programmes are not standard care [@Mohammad2013; @Jayaraman2014; @Kadhum2020].

Large scale cluster randomised trials can be complex and challenging, so before embarking on a full scale trial, we conducted a pilot study with the aim to assess the feasibility of a cluster randomised controlled trial comparing the effects of ATLS and PTC with standard care on outcomes in adult trauma patients.

>>>>

Trauma, defined as a clinical entity composed of physical injury and the associated response of the body, causes 4.3 million deaths every year [@gbd_2021_collaborators_injuries_2024]. Several trauma life support training programmes have been developed to improve the early management of patients in hospitals by providing a structured framework for assessment and treatment [@Mohammad2013; @Jayaraman2014; @Kadhum2020; @Jin2021; @putra_impact_2023; @mciver_effect_2024].

@@

There are no controlled trials of the effect of ATLS or PTC on patient outcomes, and no trials evaluating any trauma life support training programme in a general trauma population [@Mohammad2013; @Jayaraman2014; @Kadhum2020; @Jin2021; @putra_impact_2023; @mciver_effect_2024]. However, a recent cluster randomised trial found that the rural trauma team development course reduced mortality in patients with motorcycle injuries [@lule_impact_2025].

Systematic reviews call for controlled trials of ATLS and PTC [@Mohammad2013; @Jayaraman2014; @Kadhum2020], but large scale cluster randomised trials can be complex. We therefore conducted a pilot study with the aim to assess the feasibility of a cluster randomised controlled trial comparing the effects of ATLS and PTC with standard care on outcomes in adult trauma patients.
